---
title: "Reading and manipulating CAI and CUT tables"
author: "CCDM Programming Club"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette documents our workflow for basic processing of the cai data.
Including reading data, summary statistics, and visualisation.

Our package comes with a single example CAI file from _Leptosphaeria maculans_.
We know that _L. mac_ has strict compartmentalisation of some of it's genomic regions.
E.g AT-rich and RIPped regions.

First, lets load all of the packages we'll be using.

```{r}
library("codonfriend")
library("ggplot2")
library("dplyr")
library("readr")
library("magrittr")
```

We'll also set some base directories that we can use to easily find and save our data.
This allows us to switch between test and real data very easily.


```{r}
# This command finds the path on your computer.
#base_dir <- system.file("extdata", package = "codonfriend")
base_dir <- "/run/media/darcyabjones/scratch/process_fungal_genomes/cais"

out_dir <- "basics_results"
dir.create(out_dir, showWarnings = FALSE)
```

For starters, let's look at a single example file in the package.

```{r}
# Read a single cai file.
lmac_cai <- read_cai(system.file("extdata", "lmac.cai", package = "codonfriend"))
head(lmac_cai)
```

Let's look at the data.
I'm expecting sort of a bimodal distribution, e.g. one for AT-rich and one for core regions.

```{r}
ggplot(lmac_cai, aes(x = cai)) +
  geom_density()
```

This is the smoothed density of all CAI values for _L. maculans_, and it certainly isn't bimodal.
It almost appears to be a skewed normal distribution.
Let's test for normality.

```{r}
# Annoyingly, max number shapiro test can take is 5000.
# So we randomly sample 5000 cai values.
shapiro.test(sample(lmac_cai$cai, 5000))
```

Because the p-value is very low we conclude that this isn't normally distributed.
$H_0$ for shapiro-wilk test is that the data are normally distributed.

Do we see the same trend for all of the CAI files?
Let's parse them all now using `codonfriends` convenience function, `read_many`.

```{r}
cai_files <- Sys.glob(file.path(base_dir, "*.cai"))
cais <- read_many(cai_files, FUN=read_cai, colname="file") %>%
  mutate(file = gsub(".cai", "", file)) %>%
  readr::write_tsv(file.path(out_dir, "joined_cais.tsv"))
head(cais)
```

Here we've parsed all of the `.cai` files in our base directory.
What does the distribution look like for these?

```{r fig.height=4, fig.width=6}
ggplot(cais, aes(x = cai, colour = file)) +
  geom_density() +
  theme(legend.position="none")
```


So most genomes have a skewed distribution of CAIs.
There are too many genomes in here to pick out individuals, but it seems that some tend to have tighter distributions with modes closer to 1.
It's interesting to me that some of them have modes around 6-7.
Probably the genomes with wide distributions and lower modes are the ones with compartmentalisation.

What's the overall distribution of the CAIs?
I.E. not grouped by genome?

```{r fig.height=4, fig.width=6}
ggplot(cais, aes(x = cai)) +
  geom_density() +
  theme(legend.position = "none")
```

OK. So generally there is a left skew, with a mode above 7.5.
It's interesting that we don't really seem to see bimodality in the distributions, though of course it's hard to tell from such a busy plot.


## Summarising distributions

Let's look at comparing distributions of CAIs using summary statistics.

```{r}
summarised_cais <- cais %>%
  mutate(file = gsub(".cai", "", file)) %>%
  group_by(file) %>%
  summarise(
    mean = mean(cai, na.rm = T),
    median = median(cai, na.rm = T),
    q1 = quantile(cai, 0.05),
    q3 = quantile(cai, 0.95),
    sd = sd(cai, na.rm = T),
    mode = {dd <- density(cai); dd$x[which.max(dd$y)]},
    n = n()
  ) %>%
  readr::write_tsv(file.path(out_dir, "summarised_cais.tsv"))

head(summarised_cais)
```

What do the distributions of summary statistics look like?

```{r}
lapply(
  c("mean", "median", "q1", "q3", "sd", "mode", "n"),
  FUN = function(stat) {
    ggplot(summarised_cais, aes_string(x = stat)) +
      geom_density() +
      ggtitle(stat)
  }
)
```


There are quite a lot of genomes with very few genes in them.
These will just be low quality assemblies or weird annotations.
Do they affect the CAI distributions?

```{r}
ggplot(summarised_cais, aes(x = median, y = n)) + geom_point()
```

It doesn't seem like there is any bias in their central statistics.